This is an example of article title

### 2019-06-22 paper/paper.tex
note on uint32 and floats

Processors offering SIMD instructions and other forms of instruction-level parallelism started to become widely available in the early 2000s. One example is the introduction of SSE instructions in the x86 architecture~\cite[Chapter 1]{kusswurm18}. The current availability of processors with 512 bit registers, along with modern GPUs, underscore the demand for software-building tools that support vector operations. One programming language feature relevant to that goal is polymorphism, which allows programmers to abstract over data types and vectorization. Another is {\em metaprogramming}, the ability of writing programs that manipulate or generate other programs~\cite{Abelson1996}\cite[Ch.25]{JuliaDocs}. Metaprogramming enables code generation for complex algorithms customized to problem parameters and platform resources.

The interest in sorting and the great relevance of cache memory and parallelism to modern computing eventually led to research on sorting algorithms that account for these factors, following research in other topics where SIMD is more clearly beneficial. At the time SIMD CPUs became widely available there was already a trend to employ specialized hardware like GPUs for scientific applications~\cite{larsen2001fast,DBLP:conf/micro/ThompsonHO02}. Although SIMD was used mostly for multimedia applications at first~\cite{CHEN2006509,DBLP:journals/mam/SlingerlandS05}, some early alternative uses include 3D graphics~\cite{DBLP:conf/pcm/MaY02} and machine learning~\cite{DBLP:conf/europar/StreyB01}.

Following earlier work on parallel sorting in other architectures, the late 2000s saw the first practical demonstrations of sorting with modern consumer SIMD chips~\cite{DBLP:conf/IEEEpact/InoueMKN07,DBLP:journals/pvldb/ChhuganiNLMHCBKD08}. Most of the techniques utilized then are still employed in more recent works~\cite{DBLP:journals/pvldb/BalkesenATO13,DBLP:journals/pvldb/InoueT15}.

@@

The way \chipsort  works mostly follows~\cite{DBLP:journals/pvldb/InoueT15}, although \chipsort does not support yet registers with key and payload. Another main difference is that Combsort is always employed with an Insertion sort finalization.

@@

>>>>

Processors offering SIMD or other forms of instruction-level parallelism started to become widely available in the early 2000s. One example is the introduction of SSE instructions in the x86 architecture~\cite[Chapter 1]{kusswurm18}. The current availability of processors with 512 bit registers, along with modern GPUs, underscore the demand for software-building tools that support vector operations. One programming language feature relevant to that goal is polymorphism, allowing programmers to abstract over data types and vectors. Another is {\em metaprogramming}, the ability of writing programs that manipulate or generate other programs~\cite{Abelson1996}\cite[Ch.25]{JuliaDocs}. Metaprogramming enables code generation for complex algorithms customized to problem parameters and platform resources.

The interest in sorting and the great relevance of cache memory and parallelism to modern computing eventually led to research on sorting algorithms that account for these factors, following research in other topics where SIMD is more clearly beneficial. At the time SIMD CPUs became widely available there was already a trend to employ specialized hardware like GPUs for scientific applications~\cite{larsen2001fast,DBLP:conf/micro/ThompsonHO02}. Although SIMD was used mostly for multimedia applications at first~\cite{CHEN2006509,DBLP:journals/mam/SlingerlandS05}, some early alternative uses include 3D graphics~\cite{DBLP:conf/pcm/MaY02} and machine learning~\cite{DBLP:conf/europar/StreyB01}. Following earlier work on parallel sorting in other architectures, the late 2000s saw the first practical demonstrations of sorting with modern consumer SIMD chips~\cite{DBLP:conf/IEEEpact/InoueMKN07,DBLP:journals/pvldb/ChhuganiNLMHCBKD08}. Most of the techniques utilized then are still employed in more recent works~\cite{DBLP:journals/pvldb/BalkesenATO13,DBLP:journals/pvldb/InoueT15}.

@@

The way \chipsort works mostly follows~\cite{DBLP:journals/pvldb/InoueT15}, except it does not yet support registers with key and payload and there are no measures to guarantee a stable sort, such as ensuring unique keys or performing a post-processing to enforce the original order of colliding keys. Another main difference is that \chipsort always employs Combsort with an Insertion sort finalization.

@@

While our experiments focused on 32-bit integer values, similar results were observed with other types, although with possibly smaller speedups with 64-bits. Further details are omitted for the sake of brevity. For floating-point numbers, while in theory every method should perform comparably to integers, issues can appear in practice due to peculiarities such as {\tt NaN}s. This interferes with the ability of the compiler to optimize {\tt chipsort!}, although the other methods with explicit vectorization are less affected. There is ongoing work in Julia and LLVM to improve the performance of floating-point comparisons~\cite{llvmD62272,llvm3acc4236b871,Julia31442}.
